---
title: "<font color='#41CBBF'  size='5'>Dan_Danr</font>"
author: "<font color='#A8A8A8' size='3'>Carmen Samedi</font>"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 6, fig.height = 4, fig.align = "center")
```

## Library and object loading
### { .tabset}
#### Load the libraries
```{r message=FALSE, warning=FALSE}
library(arrow)
library(ggplot2)
library(dplyr)
library(plotly)
library(AnnotationHub)
library(GenomicRanges)
library(cisTopic)
library(pheatmap)
library(Seurat)
library(Signac)
library(readxl)
library(patchwork)
library(cowplot)
library(topGO)
library(org.Dm.eg.db)
library(AnnotationDbi)
library(SingleR)
library(celldex)
```

#### Load cisTopic object
```{r }
cisTopicObj <- readRDS("GSE163697_cisTopicObject_OL.Rds")

selectedModel <- cisTopicObj@selected.model

names(cisTopicObj@cell.data)

cisTopicObj <- runUmap(cisTopicObj, 
                       target = "cell", 
                       seed = 123, 
                       n_neighbors = 30, 
                       min_dist = 0.3)
```

## Plotting
### { .tabset}
#### UMAP according to accessibility 
```{r message=FALSE, warning=FALSE}
p_ncounts <- plotFeatures(cisTopicObj, 
                          target = "cell", 
                          method = "Umap", 
                          colorBy = "nAcc") 

p_ncounts + theme(legend.text = element_text(size=12),
            legend.title = element_text(size = 14),
            legend.key.width = unit(1.5, "cm") )
```

#### UMAP according to number of counts 
```{r message=FALSE, warning=FALSE}
p_ncounts <- plotFeatures(cisTopicObj, 
                          target = "cell", 
                          method = "Umap", 
                          colorBy = "nCounts") 

p_ncounts + theme(legend.text = element_text(size=12),
            legend.title = element_text(size = 14),
            legend.key.width = unit(1.5, "cm") )
```

#### UMAP according to the samples
```{r}
cisTopicObj@cell.data$sample_id <- sub("\\..*", "", cisTopicObj@cell.names)
table(cisTopicObj@cell.data$sample_id)

p_samples <- plotFeatures(cisTopicObj,
                          target = "cell",
                          method = "Umap",
                          colorBy = "sample_id")

p_samples + theme(legend.text = element_text(size=12),
            legend.title = element_text(size = 14),
            legend.key.width = unit(1.5, "cm") )
```

## Correlation matrix and a Heatmap 
```{r }
topic_mat <- modelMatSelection(cisTopicObj, target = "cell", method = "Probability")
topic_matrix_t <- t(topic_mat)

cor_matrix_topics <- cor(topic_matrix_t)
pheatmap(cor_matrix_topics[1:30, 1:30])
```

## Dan and Danr Analysis 
### { .tabset}
#### Dan and Danr coordinates from Flybase.org
```{r }
dan_chr <- "chr3R"
dan_tss <- 25184383
danr_tss <- 25139800
window <- 10000
region_coords <- cisTopicObj@region.names # Associate to region_coords all of the accessible dna
```

#### Parse and create dataframe for easy usage 
```{r}
parse_regions <- function(region_strings) {
  data.frame(
    chr = sub(":.*", "", region_strings),
    start = as.numeric(sub(".*:(\\d+)-.*", "\\1", region_strings)),
    end = as.numeric(sub(".*-(\\d+)", "\\1", region_strings)),
    stringsAsFactors = FALSE
  )
}

region_df <- parse_regions(region_coords)
``` 

#### Find regions near dan and danr 
```{r}
target_regions <- region_coords[
  (region_df$chr == dan_chr) &
    (
      (region_df$start >= dan_tss - window & region_df$end <= dan_tss + window) |
        (region_df$start >= danr_tss - window & region_df$end <= danr_tss + window)
    )
]

target_regions_gr <- GRanges(
  seqnames = sub(":.*", "", target_regions),
  ranges = IRanges(
    start = as.numeric(sub(".*:(\\d+)-.*", "\\1", target_regions)),
    end = as.numeric(sub(".*-(\\d+)", "\\1", target_regions))
  )
)

regionScores <- modelMatSelection(cisTopicObj, target = "region", method = "Z-score")
region_names <- colnames(regionScores)

# Convertir a GRanges para comparar overlaps between sets of regions 
region_gr <- GRanges(
  seqnames = sub(":.*", "", region_names),
  ranges = IRanges(
    start = as.numeric(sub(".*:(\\d+)-.*", "\\1", region_names)),
    end = as.numeric(sub(".*-(\\d+)", "\\1", region_names))
  )
)

``` 

#### Find regions that are near to dan and danr and that are in the model
```{r}
hits <- findOverlaps(target_regions_gr, region_gr)
matched_region_names <- region_names[subjectHits(hits)]
length(matched_region_names)
dan_danr_scores <- regionScores[, matched_region_names, drop = FALSE] # z-score to know how strong this region is associated to each topic 

```

## Plot the results 
### { .tabset}
#### Box plot of the z-scores of regions near dan and danr
```{r}
boxplot(dan_danr_scores, main = "Z-scores of regions near dan/danr", ylab = "Z-score")
```

#### Box plot of the topics that are active in regions near Dan and danr 
```{r}
boxplot(t(dan_danr_scores), las = 2, main = "Active topics in regions near dan/danr", ylab = "Z-score")
``` 


## Find the names of the Topics 
### { .tabset}
#### Extract the region names
```{r}
regionScores_t <- t(regionScores)
regionScores_t <- as.data.frame(regionScores_t)
regionScores_t$region <- rownames(regionScores_t)
``` 

#### Calculate the z-scores from the valid regions and pick the top indices 
```{r}
valid_regions <- intersect(rownames(regionScores_t), matched_region_names)
subset_scores <- regionScores_t[valid_regions, 1:(ncol(regionScores_t) - 1)]
topic_means <- colMeans(subset_scores)
top_topic_indices <- order(topic_means, decreasing = TRUE)[1:5]
top_topic_names <- names(topic_means)[top_topic_indices]
print(top_topic_names)
``` 

## Plotting the topics near our genes of interest 
### { .tabset}
#### UMAP of Topic 25
```{r}
p25 <- plotFeatures(
  cisTopicObj,
  method = "Umap",
  target = "cell",
  topic_contr = "Probability",
  topics = "Topic25"
)
``` 

#### UMAP of Topic 32
```{r}
p32 <- plotFeatures(
  cisTopicObj,
  method = "Umap",
  target = "cell",
  topic_contr = "Probability",
  topics = "Topic32"
)

``` 

#### UMAP of Topic 67
```{r}
p67 <- plotFeatures(
  cisTopicObj,
  method = "Umap",
  target = "cell",
  topic_contr = "Probability",
  topics = "Topic67"
)

``` 

#### UMAP of Topic 116
```{r}
p116 <- plotFeatures(
  cisTopicObj,
  method = "Umap",
  target = "cell",
  topic_contr = "Probability",
  topics = "Topic116"
)


``` 

#### UMAP of Topic 175
```{r}
p175 <- plotFeatures(
  cisTopicObj,
  method = "Umap",
  target = "cell",
  topic_contr = "Probability",
  topics = "Topic175"
)

``` 


## Seurat Analysis of the Cistopic
### { .tabset}
#### Clusters of the umap 
```{r message=FALSE, warning=FALSE}
set.seed(233)
topic_mat <- modelMatSelection(cisTopicObj, target = "cell", method = "Probability")

seurat_obj <- CreateSeuratObject(counts = topic_mat, assay = "topics")
seurat_obj <- RunTFIDF(seurat_obj)
seurat_obj <- FindTopFeatures(seurat_obj, min.cutoff= "q0")
seurat_obj <- RunSVD(seurat_obj)
seurat_obj <- RunUMAP(seurat_obj, reduction = "lsi", dims = 2:30, reduction.name = "umap.atac", reduction.key = "atacUMAP_")
seurat_obj <- FindNeighbors(seurat_obj, reduction = "lsi", dims = 2:30)
seurat_obj <- FindClusters(seurat_obj, resolution = 0.8)
DimPlot(seurat_obj, group.by = "seurat_clusters", label = TRUE)
``` 

#### Topic 32
```{r}
FeaturePlot(seurat_obj, features = "Topic32", reduction = "umap.atac", label = T)
``` 

#### Topic 25
```{r}
FeaturePlot(seurat_obj, features = "Topic25", reduction = "umap.atac", label = T)
``` 

#### Topic 67 
```{r}
FeaturePlot(seurat_obj, features = "Topic67", reduction = "umap.atac", label = T)
``` 

#### Topic 116
```{r}
FeaturePlot(seurat_obj, features = "Topic116", reduction = "umap.atac", label = T)
```

#### Topic 175 
```{r}
FeaturePlot(seurat_obj, features = "Topic175", reduction = "umap.atac", label = T)
``` 
